---
title: "HW 1 -- R STARTER"
format:
  html:
    embed-resources: true
    toc: true
    toc-location: left
    df-print: kable
execute: 
  warning: false
  message: false
---

## Simulation 1 

### a) Population distribution

Let $X \sim \text{Gamma}(\alpha = 4, \beta = 1/2)$.

```{r}
x <- seq(from = 0, to = 100, by = 0.01)
plot(x = x, y = dgamma(x, shape = 4, rate = 1/2), type = "l", xlab = "x", ylab = "f(x)", main = "Gamma population distribution")

```

### b and c) Generate random samples and calculate statistic

Let $Y = \text{median}\{X_1, \ldots, X_{30}\}$.

```{r}
# initialize our holding vector again
x_median <- c()
for (i in 1:10000) {
  
  # i is the index, the for loop runs for each value of the index
  
  # generate data
  data_sample <- rgamma(n = 30, shape = 4, rate = 1/2)
  
  # record sample median
  x_median[i] <- median(data_sample)
  
  # once both of these are done, the for loop goes to the next index value and performs the loop again
}

```

### d) Plot simulated sampling distribution of $Y = \text{median}\{X_1, \ldots, X_{30}\}$

```{r}
# plot sample median
hist(x = x_median, freq = FALSE, xlab = "Sample median", main = "Sampling distribution from Gamma population")

```

Sampling distribution seems to be centered at the population median.

### e) Calculate the mean and standard deviation of the simulated sampling distribution

Calculate $\hat{\mu}_Y$ and $\hat{\sigma}_Y$.

- We don't know the theoretical expected value and variance of the statistic, so can't compare the simulated results to anything.

```{r}
s = sd(x_median)
m = mean(x_median) 
```

### f) Estimate the probability the sample statistic is within two standard deviations of its mean

Calculate $P(\hat{\mu}_Y - 2 \hat{\sigma}_Y < Y < \hat{\mu}_Y + 2 \hat{\sigma}_Y)$.

```{r}
# P(X < 1.5)
mean(x_median < m+2*s)-mean(x_median < m-2*s)

```

## Simulation 2

### a) Population distribution

Let $X \sim \text{??}(??)$.

```{r}
x <- seq(from = 0, to = 25, by = 0.01)
plot(x = x, y = dchisq(x, df=5, ncp=0, log=FALSE), type = "l", xlab = "x", ylab = "f(x)", main = "ChiSquare population distribution")

```

### b and c) Generate random samples and calculate statistic

Let $Y = \text{??}\{X_1, \ldots, X_{30}\}$.

```{r}
# initialize our holding vector again
x_median <- c()
for (i in 1:10000) {
  
  # i is the index, the for loop runs for each value of the index
  
  # generate data
  data_sample <- rchisq(n=30, df=5, ncp=0)
  
  # record sample median
  x_median[i] <- median(data_sample)
  
  # once both of these are done, the for loop goes to the next index value and performs the loop again
}


```

### d) Plot simulated sampling distribution of $Y = \text{max}\{X_1, \ldots, X_{30}\}$

```{r}
# plot sample median
hist(x = x_median, freq = FALSE, xlab = "Sample median", main = "Sampling distribution from ChiSquare population")


```

### e) Calculate the mean and standard deviation of the simulated sampling distribution

Calculate $\hat{\mu}_Y$ and $\hat{\sigma}_Y$.

```{r}
s = sd(x_median)
m = mean(x_median) 


```

### f) Estimate the probability the sample statistic is within two standard deviations of its mean

Calculate $P(\hat{\mu}_Y - 2 \hat{\sigma}_Y < Y < \hat{\mu}_Y + 2 \hat{\sigma}_Y)$.

```{r}
# P(X < 1.5)
mean(x_median < m+2*s)-mean(x_median < m-2*s)


```